---
title: "Trust in Science: Global Patterns & Change"
format:
  dashboard:
    logo: images/Wellcome_Trust_logo.svg
    orientation: rows
    nav-buttons: [linkedin, github]
    linkedin: www.linkedin.com/in/jacob-u-perez
    github: https://github.com/jacobperez-debug
    theme: journal
server: shiny
editor: visual
---

```{r}
#| label: libraries
#| context: setup

library(ggplot2)
library(shiny)
library(shinydashboard)
library(tidyverse)
```

```{r}
#| label: read in data
#| context: setup

trust2018 <- read.csv(here::here("trust2018.csv"))
trust2020 <- read.csv(here::here("trust2020.csv"))
trust_change <- read.csv(here::here("trustchange.csv"))
df_wgm2018 <- read.csv(here::here("wgm2018.csv"))
df_wgm2020 <- read.csv(here::here("wgm2020.csv"))
```

# Trust Across Regions {.tabset}

## inputs {.sidebar}

```{r}
checkboxGroupInput("region_select", "Choose Regions:",
                   choices = c("Africa", "Americas", "Asia", "Europe", 
                               "Middle East and North Africa", "Russia/Caucasus/Central Asia"),
                   selected = c("Africa", "Americas", "Asia", "Europe", 
                               "Middle East and North Africa", "Russia/Caucasus/Central Asia"))
```

## Global Trust in Scientists Plots (Row 2) {height = 70%}

```{r}
plotOutput('plot.2018')
```

```{r}
plotOutput('plot.2020')
```

## Global Trust in Scientists Cards (Row 3)

### Card 1 {orientation = "columns"}

#### Facts {orientation = "columns}

```{r}
trusted_by_2018 <- df_wgm2018 %>%
  filter(!is.na(Q11C)) %>%
  mutate(
    trusted = if_else(Q11C %in% c(1, 2), 1, 0),
    weighted_trusted = wgt * trusted
  ) %>%
  group_by(region) %>%
  summarise(
    percent_trusted = 100 * sum(weighted_trusted, na.rm = TRUE) / sum(wgt, na.rm = TRUE)
  ) %>%
  mutate(
    Percent = round(percent_trusted, 2),
    Region = region
  ) %>%
  select(Region, Percent)

knitr::kable(trusted_by_2018)
```

### Interactive Cards

```{r}
valueBoxOutput("avg.2018")
```

```{r}
valueBoxOutput("avg.2020")
```

### Card 3

```{r}
trusted_by_2020 <- df_wgm2020 %>%
  filter(!is.na(W5C)) %>%
  mutate(
    trusted = if_else(W5C %in% c(1, 2), 1, 0),
    weighted_trusted = WGT * trusted
  ) %>%
  group_by(region) %>%
  summarise(
    percent_trusted = 100 * sum(weighted_trusted, na.rm = TRUE) / sum(WGT, na.rm = TRUE)
  ) %>%
  mutate(
    Percent = round(percent_trusted, 2),
    Region = region
  ) %>%
  select(Region, Percent)

knitr::kable(trusted_by_2020)
```

# Country Trends Over Time

## labels {.sidebar} 

```{r}
sliderInput("change_range", "Select Trust Change Range (%):",
            min = -40, max = 40, value = c(-10, 10))
```

## Row 2

```{r}
plotOutput("change_lollipop")
```

## Country Trust Tables and Cards (Row 3) {height = 30%}

### Country Tables {orientation = "column"}

```{r}
#| message: false

top5 <- trust_change %>%
  arrange(desc(change)) %>%
  head(5) %>%
  mutate(
    Country = country_name,
    "Percent Change" = change
    ) %>%
  select(Country, `Percent Change`)

knitr::kable(top5, digits = 2)
```

```{r}
#| message: false

bot5 <- trust_change %>%
  arrange(change) %>%
  select(country_name, change) %>%
  head(5) %>%
  mutate(
    Country = country_name,
    "Percent Change" = change
    ) %>%
  select(Country, `Percent Change`)

knitr::kable(bot5, digits = 2)
```

### Biggest Decrease

```{r}
valueBoxOutput("biggest_decrease")
```

### Biggest Increase

```{r}
valueBoxOutput("biggest_increase")
```

```{r}
#| context: server

filtered_2018 <- reactive({
  trust2018 %>%
    filter(region %in% input$region_select)
})

output$plot.2018 <- renderPlot({
  ggplot(
    filtered_2018(), 
    aes(x = percent)) +
    geom_density(fill = "#65a345", alpha = .5) +
    labs(title = "Trust in Scientists (2018)", 
         x = "% Trust",
         y = "") +
    theme_minimal()
})

output$avg.2018 <- renderValueBox({
  avg_2018 <- mean(filtered_2018()$percent, na.rm = TRUE)
  
  valueBox(
    value = paste0(round(avg_2018, 1), "%"),
    subtitle = "Avg Trust (2018)",
    icon = icon("calendar"),
    color = "orange"
  )
})

filtered_2020 <- reactive({
  trust2020 %>%
    filter(region %in% input$region_select)
})

output$plot.2020 <- renderPlot({
  ggplot(
    filtered_2020(), 
    aes(x = percent)) +
    geom_density(fill = "#65a345", alpha = .5) +
    labs(title = "Trust in Scientists (2020)", 
         x = "% Trust",
         y = "") +
    theme_minimal()
})

output$avg.2020 <- renderValueBox({
  avg_2020 <- mean(filtered_2020()$percent, na.rm = TRUE)
  
  valueBox(
    value = paste0(round(avg_2020, 1), "%"),
    subtitle = "Avg Trust (2020)",
    icon = icon("calendar"),
    color = "orange"
  )
})

filtered_change <- reactive({
  trust_change %>%
    filter(change >= input$change_range[1],
           change <= input$change_range[2])
})

output$change_lollipop <- renderPlot({
  ggplot(filtered_change(), aes(x = change, y = reorder(country_name, change))) +
    geom_segment(aes(x = 0, xend = change, yend = country_name), color = "gray") +
    geom_point(color = "steelblue", size = 3) +
    labs(title = "Change in Trust (2018 to 2020)",
         x = "Change in % Trust",
         y = "") +
    theme_minimal() + 
    theme(
      axis.text.y = element_blank()
    )
})

output$biggest_increase <- renderValueBox({
  top_country <- filtered_change() %>% arrange(desc(change)) %>% head(1)
  
  if (nrow(top_country) == 0) {
    valueBox(
      value = "No countries",
      subtitle = "Highest Percentage",
      icon = icon("arrow-up"),
      color = "green"
    )
  } else {
    valueBox(
      value = paste0(as.character(top_country$country_name, ": +", round(top_country$change, 1), "%")),
      subtitle = "Highest Percentage",
      icon = icon("arrow-up"),
      color = "green"
    )
  }
})

output$biggest_decrease <- renderValueBox({
  bottom_country <- filtered_change() %>% arrange(change) %>% head(1)
  
  if (nrow(bottom_country) == 0) {
    valueBox(
      value = "No countries",
      subtitle = "Lowest Percentage",
      icon = icon("arrow-down"),
      color = "red"
    )
  } else {
    valueBox(
      value = paste0(as.character(bottom_country$country_name, ": ", round(bottom_country$change, 1), "%")),
      subtitle = "Lowest Percentage",
      icon = icon("arrow-down"),
      color = "red"
    )
  }
})
```

